Rule-Based Normalization of German Twitter Messages

### 2013-08-27 darmstadt_gscl_2013/article_twitter_normalization.tex
related work partially re-written

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{paralist} % for inparaenum
\usepackage{multirow} % for multirow
\usepackage{tabularx} % for centering in table
\usepackage{graphicx} % for \includegraphics
\usepackage{url}

@@

@@

  This article gives an overview of existing approaches to the problem
  of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  language texts. These approaches are classified with regard to the
  size of text spans and knowledge inference mechanisms which they
  rely on in their work. Additionally to that, we conduct quantitative
  and qualitative analyses of unknown words in German Twitter
  messages, in order to see how relevant the OOV and text
  normalization problems are for this particular kind of micro-texts
  and what the characteristics of these OOVs are. In a concluding
  step, we present a set of ad-hoc techniques which are supposed to
  tackle some of the most prominent disturbing effects found during
  the analyses and show how this set of techniques helps us lower the
  average rate of out-of-vocabulary tokens in Twitter messages and how
  this lower OOV-rate in turn helps improve the quality of automatic
  part-of-speech tagging.

@@

In the next Section, we will give a short overview of existing
scientific approaches to the problem of tackling out-of-vocabulary
words in non-standard texts. After that, in Section
\ref{error:analysis} we will analyze which types of noisiness
phenomena are especially characteristic for German Twitter. Section
\ref{normalization} will subsequently describe an automatic procedure
for mitigating some of the most prominent of those effects. In a
concluding step, we will perform an evaluation of the results of this
procedure and give further suggestions for future research.

@@

Before proceeding with the description of recent methods for noisy
text normalization (NTN), we first would like to define the criteria
by which these methods could be classified. It should be noted that
there already exist several classifications of NTN approaches
including, for example, Kukich (\cite{kukich}) and Kobus et
al. (\cite{kobus}).

Kukich (\cite{kukich}), for instance, divided all NTN techniques into
six classes:
\begin{enumerate}
\item minimum edit distance techniques;
\item similarity key techniques;
\item rule-based techniques;
\item \textit{n}-gram based techniques;
\item probabilistic techniques;
\item neural nets.
\end{enumerate}
Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
\textit{metaphors} and split them into the following three groups:
\begin{enumerate}
\item ``spell checking'' metaphor;
\item ``translation'' metaphor;
\item ``speech recognition'' metaphor.
\end{enumerate}

Though both of these divisions seem to be justified to some extent, it
is difficult to determine using them whether an NTN approach that
detects and restores incorrectly spelled words on the basis of
phonetical \textit{n}-gram statistics should fall into the
\textit{n}-gram based, probabilistic, spell checking or speech
recognition class.

This confusion is explained by the fact that the above classifications
both rely on several independent criteria at the same time, however
each of these criteria characterizes an NTN system from a different
point of view. As a consequence of this, unambiguous assignment of an
NTN approach to one particular class often becomes impossible. In
order to avoid this, we instead suggest using separate classifications
for each type of involved criteria.

One of such criteria which in our opinion would be worth a separate
classification is \emph{segmentation level} that is used
\begin{inparaenum}[\itshape a\upshape)]
\item to infer in-vocabulary (IV) equivalents for OOV
  tokens\label{candidate:inferring} and
\item to choose the most probable variant\label{candidate:selection}
  among multiple possible suggestions\footnote{For the cases, when
    segments of different lengths are used for tasks
    \ref{candidate:inferring} and \ref{candidate:selection}, we note
    it explicitly in our classification on which segmentation length
    each of these subtasks relies.}.
\end{inparaenum} For this criterion, we propose division into the following
classes:
\begin{enumerate}
  \item graphematic\footnote{Depending on whether phonetical
    information is involved or not at this level, this class could be
    further divided into a phonographematic and purely graphematic
    subclasses.};
  \item lexical;
  \item phrasal.
\end{enumerate}
Each broader level of this hierarchy is supposed to either incorporate
or ignore information provided by its narrower subsegments. In this
way, we only need to mention one (the broadest) hierarchical class for
the cases when multiple segmentation levels are involved by some
techniques.

The second criterion regards the \emph{type of information induction}
that is used to devise the correction rules. This leads us to the
usual NLP-taxonomy which divides all approaches into:
\begin{enumerate}
  \item rule-based;
  \item statistical\footnote{Depending on the type of training data
    required, this class is in turn usually divided into unsupervised,
    semi-supervised, and supervised groups.};
  \item and hybrid ones.
\end{enumerate}

According to these two classification criteria, recent approaches to
the NTN task could be grouped together as follows.

The earlier works on NTN commonly relied on either purely graphematic
or phonographematic levels of segmentation for deriving normalization
variants of incorrectly spelled words. To purely graphematic systems
belong methods suggested by Brill and Moore (\cite{brill}), Sproat et
al. (\cite{sproat}), and Clark (\cite{alexander-clark}). As
phonographematic approaches one could regard works done by Toutanova
and Moore (\cite{toutanova}), Choudhury et al. (\cite{choudhury}),
Cook and Stevenson (\cite{cook}), Han and Baldwin (\cite{han})
etc. With regard to the type of information inference, most of these
methods were supervised with the exceptions of Cook and Stevenson
(\cite{cook}) and Han and Baldwin (\cite{han}) who claimed to use an
unsupervised technique. Sproat et al. (\cite{sproat}) described in
their article both a supervised and unsupervised approach.

Starting from the second half of the 2000s, the raising influence and
improved quality of machine translation tools led to the development
of NTN technologies which used broader levels of segmentation. In
\cite{aw}, Aw et al. suggested a supervised statistical system for
normalization of SMS-messages which operated on automatically aligned
phrases A few years later, Clark and Araki (\cite{clark-araki})
described a purely rule-based method which used mappings of
non-standard words and phrases to their corresponding standard
language forms.

As noted by Kobus et al. (\cite{kobus}), NTN methods relying on either
graphematic or phrasal segments usually revealed complementary
strengths and weaknesses. This notion led NLP scientists to the idea
that by incorporating multiple levels of the language into one NTN
system the total performace of the whole system would improve as
different sources of information would benefit from each other. As a
consequence of this, a wealth of combined techniques emerged in the
past few years. Among these we should especially mention works by
Kobus et al. (\cite{kobus}), Kaufmann (\cite{kaufmann}), and Oliva et
al. (\cite{oliva}). The majority of these systems used the whole range
of segmentation levels from phonographematatic to phrasal one, and in
many cases they also applied different knowledge inference mechanisms
to different levels of the language.

@@

It should however be noted that almost all of the above methods mainly
concentrated on only English data. A few exceptions from that are
approaches suggested by Beaufort et al. (\cite{beaufort}) for French,
and Oliva et al. (\cite{oliva}) for Spanish. To find out which
peculiarities of ill-formed words are characteristic for German, we
will perform a quantitive and qualitative analysis of unknown words in
German Twitter in the next Section in order to see what kind of NTN
techniques would be most suitable for handling such words there.

\section{Analysis of Unknown Tokens}\label{error:analysis}

@@

We classified found OOV tokens into the following three groups
according to the reasons why these tokens could have been omitted from
corresponding applications' dictionaries:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionaries
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or did not exist at the time when
    dictionaries were created. Another reason for the inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication;\label{style}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    group, we considered incorrect spellings of words encountered in
    text.\label{spell}
\end{enumerate}

@@

@@

@@

  McVeigh, T.: %
  Text messaging turns 20. %
  The Observer (December 1, 2012)

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{paralist} % for inparaenum
\usepackage{multirow} % for multirow
\usepackage{tabularx} % for centering in table
\usepackage{graphicx} % for \includegraphics
\usepackage{url}
\usepackage{amsmath}

@@

\DeclareMathOperator*{\argmax}{arg\,max}

@@

  %% This article gives an overview of existing approaches to the problem
  %% of out-of-vocabulary (OOV) tokens and noisiness phenomena in natural
  %% language texts. These approaches are classified with regard to the
  %% size of text spans and knowledge inference mechanisms which they
  %% rely on in their work. Additionally to that, we conduct quantitative
  %% and qualitative analyses of unknown words in German Twitter
  %% messages, in order to see how relevant the OOV and text
  %% normalization problems are for this particular kind of micro-texts
  %% and what the characteristics of these OOVs are. In a concluding
  %% step, we present a set of ad-hoc techniques which are supposed to
  %% tackle some of the most prominent disturbing effects found during
  %% the analyses and show how this set of techniques helps us lower the
  %% average rate of out-of-vocabulary tokens in Twitter messages and how
  %% this lower OOV-rate in turn helps improve the quality of automatic
  %% part-of-speech tagging.

 In this article, we conduct quantitative and qualitative analyses of
 unknown words in German Twitter messages, and propose a normalization
 method which prepares German Twitter text for standard text
 processing tools. In the first part, the prevalence of different
 types of out-of-vocabulary (OOV) tokens and non-standard language in
 German Twitter data is determined. In a second step, we present a set
 of ad-hoc techniques which can tackle some of the most prominent
 effects found during the analyses. We show how this set of techniques
 helps us lower the average rate of out-of-vocabulary tokens in
 Twitter messages and how this lower OOV-rate in turn helps improve
 the quality of automatic part-of-speech tagging.

@@

In the next Section, we give a short overview of existing scientific
approaches to the problem of tackling out-of-vocabulary words in
non-standard texts. After that, in Section \ref{error:analysis} we
will analyze which types of noisiness phenomena are especially
characteristic for German Twitter. Section \ref{normalization} will
subsequently describe an automatic procedure for mitigating some of
the most prominent of those effects. In a concluding step, we will
perform an evaluation of the results of this procedure and give
further suggestions for future research.

@@

%% Before proceeding with the description of recent methods for noisy
%% text normalization (NTN), we first would like to define the criteria
%% by which these methods could be classified. It should be noted that
%% there already exist several classifications of NTN approaches
%% including, for example, Kukich (\cite{kukich}) and Kobus et
%% al. (\cite{kobus}).

%% Kukich (\cite{kukich}), for instance, divided all NTN techniques into
%% six classes:
%% \begin{enumerate}
%% \item minimum edit distance techniques;
%% \item similarity key techniques;
%% \item rule-based techniques;
%% \item \textit{n}-gram based techniques;
%% \item probabilistic techniques;
%% \item neural nets.
%% \end{enumerate}
%% Kobus (\cite{kobus}), on the contrary, referred to NTN methods as
%% \textit{metaphors} and split them into the following three groups:
%% \begin{enumerate}
%% \item ``spell checking'' metaphor;
%% \item ``translation'' metaphor;
%% \item ``speech recognition'' metaphor.
%% \end{enumerate}

%% Though both of these divisions seem to be justified to some extent, it
%% is difficult to determine using them whether an NTN approach that
%% detects and restores incorrectly spelled words on the basis of
%% phonetical \textit{n}-gram statistics should fall into the
%% \textit{n}-gram based, probabilistic, spell checking or speech
%% recognition class.

%% This confusion is explained by the fact that the above classifications
%% both rely on several independent criteria at the same time, however
%% each of these criteria characterizes an NTN system from a different
%% point of view. As a consequence of this, unambiguous assignment of an
%% NTN approach to one particular class often becomes impossible. In
%% order to avoid this, we instead suggest using separate classifications
%% for each type of involved criteria.

%% We use two independent criteria to classify normalization methods.
%% One of such criteria which in our opinion would be worth a separate
%% classification
%% is the \emph{segmentation level} that is used
%% \begin{inparaenum}[\itshape a\upshape)]
%% \item to infer in-vocabulary (IV) equivalents for OOV
%%   tokens\label{candidate:inferring} and
%% \item to choose the most probable variant\label{candidate:selection}
%%   among multiple possible suggestions\footnote{If
%%     segments of different lengths are used for tasks
%%     \ref{candidate:inferring} and \ref{candidate:selection}, this can
%%     be noted explicitly.}.
%% \end{inparaenum} For this criterion, we propose division into the following
%% classes:
%% \begin{enumerate}
%%   \item graphematic\footnote{Depending on whether phonetic
%%     information is involved at this level, this class can be
%%     further divided into a phonographematic and purely graphematic
%%     subclasses.};
%%   \item lexical;
%%   \item phrasal.
%% \end{enumerate}
%% Each broader level of this hierarchy is supposed to either incorporate
%% or ignore information provided by its narrower subsegments. In this
%% way, we only need to mention one (the broadest) hierarchical class for
%% the cases when multiple segmentation levels are involved by some
%% techniques.

%% The second criterion regards the \emph{type of information induction}
%% that is used to devise the correction rules. This leads us to the
%% usual NLP-taxonomy which divides all approaches into:
%% \begin{enumerate}
%%   \item rule-based;
%%   \item statistical\footnote{Depending on the type of training data
%%     required, this class is in turn usually divided into unsupervised,
%%     semi-supervised, and supervised groups.};
%%   \item and hybrid ones.
%% \end{enumerate}

%% According to these two classification criteria, recent approaches to
%% the NTN task could be grouped together as follows.

The earlier works on normalization of noisy messages were greatly
influenced by text restoration techniques commonly applied in the
domains of speech and optical character recognition. One of the most
well-established techniques in those fields at the beginning of the
1990-s was the method of ``noisy channel model'' (NCM) first
introduced by Shannon in \cite{shannon}. In this method, an input text
$T$ was regarded as a distorted version of some initial clean language
signal $S$. And in order to find the original sequence, one had to
solve the equation $\mathop{arg\,max} P(S|T)$ by computing an
equivalent expression $\mathop{arg\,max} P(T|S)P(S)$. In this last
form, the term -- $P(T|S)$ -- was usually referred to as \textit{error
  model}, i.e. one which showed how likely it was that the given
non-standard form $T$ could have resulted from the assumed standard
counterpart $S$. The latter part of the equation -- $P(S)$ -- was
called the \textit{language model} and accounted for the probability
of form $S$ to appear in standard text in general. %% It is however
%% not specified by NCM, what should regarded as language model, what
%% has to be considered as an error model and eventually how a set of
%% standard language equivalents $S$ should be derived for given
%% non-standard form $T$.

Among works belonging to the NCM framework, we could mention the
approach suggested by Mayes, Damerau et al. (\cite{mayes}) who used a
uniform error model and calculated $P(S)$ on the basis of word bigrams
statistics. As a set $S$ of possible replacement suggestions, the
authors regarded the original input word $T$ as well as all words in
dictionary which could be derived from the original word by applying
at most one edit operation either insertion, or deletion, or
substituion of a single character. In \cite{church}, Church and Gale
suggested a more elaborated error model in which every edit operation
was weihgted according to statistics gathered from a training corpus
and conditioned on the letter context in which this edit operation
apeeared. Brill and Moore (\cite{Brill}) further enhanced this model
by considering edit operations not only on single letters but rather
on word subsegments of different lengths. Toutanova and Moore
(\cite{toutanova}) improved the performance of the latter work by
23~\% by incorporating phonetical information. Eventually, Alexander
Clark (\cite{clark}) was one among the first who addressed the text
noisiness problem in the genre of Internet-based communication
(IBC). In his presented system, multiple indpendent modules were used
to perform tokenization, normalization and correction of Usenet forum
posts. The system relied on a 3-gram language model and a trainable
stateless stochastic transduces as error model, however no particular
details about its implementation were provided in the article. In
\cite{choudhury}, Choudhury et al. proposed

Bangalore et al. - MT

suggested an NCM
approach in which they derived the set $S$ by applying one of the four
operations: insertion, deletion or substitution of a single character
or transposition of two adjacent characters to string $T$ and then
leaving only variants which were present in vocabulary. Original word
$T$ was also included in the generated set. As language model, the
authors used word bigrams. As error model, a uniform distribution was
used which assigned equal probabilities to all generated variants and
a constant probability $\alpha$ to the original word.

%% According to this, the main task of NCM consisted in finding such $S$
%% which could maximize the probability of $S$ given $T$. By using the
%% Bayesian rule of probability, $P(S|T)$ was defined as
%% $\frac{P(T|S)P(S)}{P(T)}$ and in order to find $arg$ and further as
%% $P(T|S)P(S)$ since the denominator $P()$

%%  commonly relied on either purely graphematic or phonographematic
%%  levels of segmentation for deriving normalization variants of
%%  incorrectly spelled words. Methods suggested by Brill and Moore
%%  (\cite{brill}), Sproat et al. (\cite{sproat}), and Clark
%%  (\cite{alexander-clark}) are purely graphematic, whereas works by
%%  Toutanova and Moore (\cite{toutanova}), Choudhury et
%%  al. (\cite{choudhury}), Cook and Stevenson (\cite{cook}), Han and
%%  Baldwin (\cite{han}), etc., can be classified as
%%  phonographematic. With regard to the type of information inference,
%%  most of these methods were supervised with the exceptions of Cook and
%%  Stevenson (\cite{cook}) and Han and Baldwin (\cite{han}) who claim to
%%  use an unsupervised technique. Sproat et al. (\cite{sproat})
%%  described in their article both a supervised and an unsupervised
%%  approach.

%% Starting from the second half of the 2000s, the raising influence and
%% improved quality of machine translation tools led to the development
%% of NTN technologies which used broader levels of segmentation. In
%% \cite{aw}, Aw et al. suggested a supervised statistical system for
%% normalization of mobile text messages which operates on automatically
%% aligned phrases. A few years later, Clark and Araki
%% (\cite{clark-araki}) described a purely rule-based method which uses
%% mappings of non-standard words and phrases to their corresponding
%% standard language forms.

As noted by Kobus et al. (\cite{kobus}), NTN methods relying on either
graphematic or phrasal segments usually reveal complementary strengths
and weaknesses. This notion led NLP scientists to the idea that by
incorporating multiple levels of the language into one NTN system the
total performace of the whole system would improve as different
sources of information would benefit from each other. As a consequence
of this, a wealth of combined techniques emerged in the past few
years. Among these we should especially mention works by Kobus et
al. (\cite{kobus}), Kaufmann (\cite{kaufmann}), and Oliva et
al. (\cite{oliva}). The majority of these systems used the whole range
of segmentation levels from phonographematatic to phrasal, and in many
cases they also applied different knowledge inference mechanisms to
different levels of the language.
% % *TS*
% It's not quite clear what these approaches actually DO -  for
% example the combined methods above or some of the early
% approaches. I liked how "Clark and Araki" was described in half a
% sentence, maybe add something like that for one or two of the papers
% in the first paragraph ("earlier works") and also for the last
% paragraph ("combined methods"). I can shorten the "combined methods"
% paragraph more if needed.
%

@@

It should however be noted that almost all of the above methods mainly
concentrated on only English data. A few exceptions are the
approaches suggested by Beaufort et al. (\cite{beaufort}) for French,
and Oliva et al. (\cite{oliva}) for Spanish. Since the peculiarities
of short messages are often language-specific, we
perform a quantitive and qualitative analysis of unknown words in
German Twitter in the next Section in order to see what kind of NTN
techniques would be most suitable for handling such words there.
\section{Analysis of Unknown Tokens}\label{error:analysis}

@@

We classified found OOV tokens into the following three groups
according to the reasons why these tokens could have been omitted from
corresponding applications' dictionaries:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionaries
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or did not exist at the time when
    dictionaries were created. Another reason for the inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication;\label{style}
  \item \textbf{Misspellings}. In the scope of this group, we
    considered incorrect spellings of words encountered in
    text.\label{spell}
\end{enumerate}

@@

\bibitem[1991]{church}
  Church, K., Gale, W.: %
  Probability scoring for spelling correction. %
  Statistics and Computing. {\bfseries 1} %
  (1991) 93--103 %

\bibitem[2003]{clark}
  Clark, A.,: %
  Pre-Processing very noisy text. %
  In Proceedings of Workshop on Shallow Processing of Large Corpora. %
  Corpus Linguistics. %
  Lancaster (2003)

@@

\bibitem[1991]{mayes}
  Mayes, E., F. Damerau, et al.: %
  Context Based Spelling Correction. %
  Information Processing and Management. 27(5) %
  (1991) 517--522

@@

@@

\bibitem[1948]{shannon}
  Shannon, C. E.: %
  A mathematical theory of communication. %
  Bell system techical journal, 27:379--423 %
  (1948) 623--656
